---
title: "andrewBowen_R_Week2Discussion"
author: "Andrew bowen"
date: "2022-07-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question

I'm a big NBA/sports fan. I'm curious to see which players in the NBA are most productive. To simplify the example, I'm going to use the pre-game counting stats of points, assists, and rebounds to see the top 10 players and bottom 10 players in each category.


## Pulling Data
I pulled 2021-22 NBA season data from [basketball-reference.com](https://www.basketball-reference.com/leagues/NBA_2022_per_game.html).

```{r}
df = read.csv('./data/nba-stats-21-22.csv', sep=',', header=TRUE)

head(df, 15)
```

```{r}
summary.data.frame(df)
```

Getting the top 10 players in points, assists, and rebounds, respectively. We only want certain columns from our `data.frame`, so we'll 
```{r}
# Sorting the data frame on total assist column, in decreasing order
ast_sorted <- df[order(df$AST[df$Tm != ""], decreasing=TRUE), c("Player", "Tm", "AST")]

head(ast_sorted, 10)
```
Check out those players that didn't involve their teammates
```{r}
tail(ast_sorted, 10)
```



Checking out top 10 and bottom 10 of points
```{r}
pts_sorted <- df[order(df$PTS, decreasing=TRUE), c("Player", "PTS")]

head(pts_sorted, 10)
```

```{r}
tail(pts_sorted, 10)
```


Grabbing rebounds. (Get it?)
```{r}
# Using total rebounds (TRB column name) to include offensive and defensive rebs
rbs_sorted <- df[order(df$TRB, decreasing=TRUE), c("Player", "TRB")]

head(rbs_sorted, 10)
```
```{r}
tail(rbs_sorted, 10)

```
## Getting per-game averages
```{r}
df$APG <- df$AST / df$G
df$PPG <- df$PTS / df$G
df$RPG <- df$TRB / df$G


head(df, 15)
```


## Plotting our data using `ggplot2`
```{r}
library(ggplot2)

# Plotting points per game vs assists per game 
ggplot(df, aes(x=APG, y=PPG))+
  geom_point(size=2, shape=23)
```



