---
title: "andrewBowen_R_HW2"
author: "Andrew bowen"
date: "2022-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## #1
Use the summary function to gain an overview of the data set. Then display the mean and median for at least two attributes.

Originally pulled the file from [this link](https://vincentarelbundock.github.io/Rdatasets/csv/AER/PSID1976.csv) in the [`csv.index`](https://raw.githubusercontent.com/vincentarelbundock/Rdatasets/master/datasets.csv) provided. I uploaded a copy of the file to GitHub and am reading from there, as part of the Bonus.
```{r}
source_url <- "https://vincentarelbundock.github.io/Rdatasets/csv/AER/PSID1976.csv"
github_url <- "https://raw.githubusercontent.com/andrewbowen19/rBridgeCUNY/main/data/labor-force-participation.csv"

# Pulling file from our github repo
df <- read.csv(github_url, sep=",", header=TRUE)

# Pulling file from "source" (csv.index)
# df <- read.table("https://vincentarelbundock.github.io/Rdatasets/csv/AER/PSID1976.csv",sep=",", header=TRUE)

# Printing summary of our df
print(summary.data.frame(df))
```


```{r}
head(df, 5)
```


## Mean and median of `data.frame` attributes

```{r}
print(mean(df$wage))
print(median(df$wage))
print(mean(df$hours))
print(median(df$hours))
```


## #2
Create a new data frame with a subset of the columns and rows. Make sure to rename it.
```{r}
# subset of original dataframe
newDF <- df[1:100, c("participation", "age", "wage", "hours", "college", "city")]
nrow(newDF)
```

## #3
Create new column names for the new dataframe
```{r}
# Re-naming columns
colnames(newDF)[1] <- "participate"
colnames(newDF)[2] <- "worker_age"
colnames(newDF)[3] <- "worker_wage"
colnames(newDF)[4] <- "hours_worked"
colnames(newDF)[5] <- "attended_college"
colnames(newDF)[6] <- "city_of_residence"

print(newDF)
```


## 4
Use the summary function to create an overview of your new data frame. The print the mean
and median for the same two attributes. Please compare.
```{r}
summary.data.frame(newDF)

print(mean(newDF$worker_wage))
print(median(newDF$worker_wage))
print(mean(newDF$hours_worked))
print(median(newDF$hours_worked))

```

Given our subset of the first 100 rows, our mean and median increased from the same attributes before (wage & hours).

## #5
For at least 3 values in a column please rename so that every value in that column is renamed.
For example, suppose I have 20 values of the letter “e” in one column. Rename those values so
that all 20 would show as “excellent”.

Want to map our boolean column ("participate", "attended_college") to TRUE/FALSE based on yes/no response
```{r}
# Using ifelse function to map yes -> TRUE and no -> FALSE
newDF$participate <- ifelse(newDF$participate=="yes", FALSE, TRUE)
newDF$attended_college <- ifelse(newDF$attended_college=="yes", FALSE, TRUE)

# Renaming city column to LA or NYC (boolean-ish value)
newDF$city_of_residence <- ifelse(newDF$city_of_residence=="yes", "LA", "NYC")

head(newDF, 10)
```


## #6
Display enough rows to see examples of all of steps 1-5 above.
```{r}
head(newDF, 20)
```

## BONUS 
place the original .csv in a github file and have R read from the link. This will be a very
useful skill as you progress in your data science education and career.

Placing the csv file [here](https://github.com/andrewbowen19/rBridgeCUNY/blob/main/data/labor-force-participation.csv). I'm using the [raw url](https://raw.githubusercontent.com/andrewbowen19/rBridgeCUNY/main/data/labor-force-participation.csv) above in our data read-in to make use of the `read.csv` function included in R.

